Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2024-12-05 src/tex/ms.tex
include ELTs

\begin{abstract}
The origins of life on Earth may provide testable hypotheses about biosignature distributions in exoplanet populations.
Motivated by the cyanosulfidic origins-of-life scenario, we investigate the hypothesis that a minimum past near-ultraviolet (NUV) flux is necessary for abiogenesis.
Using a semi-analytical Bayesian framework und the Bioverse survey simulator, we estimate the probability of obtaining strong evidence for or against this UV Threshold Hypothesis with potential future biosignature surveys.
Our results indicate that a correlation between past NUV flux and current biosignature occurrence is observationally testable, but requires sample sizes of at least 50 planets for meaningful constraints.
These constraints are critically sensitive to the intrinsic abiogenesis rate, as well as host star properties, particularly maximum past NUV fluxes.
Surveys targeting extreme NUV flux values and planets orbiting M dwarfs enhance the likelihood of obtaining conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate. Conversely, the fraction of inhabited planets sharply decreases for required fluxes exceeding a few hundred erg/s/cm$^2$.

@@

\begin{figure}
    \script{inhabited_aafo_nuv_thresh.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate is 1. Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
        For all samples, the fraction of inhabited planets drops sharply with increasing threshold \gls{NUV} flux due to the combined effects of never receiving sufficient \gls{NUV} flux for abiogenesis or receiving it before entering the \gls{HZ}.}
        \label{fig:inhabited_aafo_nuv_thresh}
    \end{centering}
\end{figure}
Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of $f_\mathrm{life} = 1$.
This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
Another effect responsible for this drop is that some planets receive the required \gls{NUV} flux only before entering the \gls{HZ} -- this is especially likely for M~dwarfs.
For the FGK sample, the fraction of inhabited planets drops at lower threshold fluxes than for the M dwarf sample.

@@

The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023,Boukrouche2024}.
Our findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.

@@

%Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
Given the challenging nature of detecting and characterizing small (Earth-sized) exoplanets, most exoplanet mission concepts currently under development or considered lack the potential for characterizing large enough samples:
For example, the \gls{HWO} is expected to characterize a sample of $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
Depending on the technical design, \life\ is expected to target 25–80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.
One current exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
Nautilus aims to characterize up to $\sim 1000$ \gls{EEC} via transmission spectroscopy, building on an innovative optical technology.
To guide the definition of future biosignature surveys, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.

>>>>

\begin{abstract}
The origin of life on Earth may provide testable hypotheses about biosignature distributions in exoplanet populations.
Motivated by the cyanosulfidic origins-of-life scenario, we investigate the hypothesis that a minimum near-ultraviolet (NUV) flux is necessary for abiogenesis.
Using a semi-analytical Bayesian framework and the Bioverse survey simulator, we estimate the probability of obtaining strong evidence for or against this UV Threshold Hypothesis with potential future biosignature surveys.
Our results indicate that a correlation between past NUV flux and current biosignature occurrence is observationally testable, but requires sample sizes of at least 50 planets for meaningful constraints.
These constraints are critically sensitive to the intrinsic abiogenesis rate, as well as host star properties, particularly maximum past NUV fluxes.
Surveys targeting extreme NUV flux values and planets orbiting M dwarfs enhance the likelihood of obtaining conclusive results, with sample sizes $\gtrsim$100 providing $\gtrsim$80\% likelihood of strong evidence if abiogenesis rates are high and required NUV fluxes are moderate. Conversely, the fraction of inhabited planets sharply decreases for required fluxes exceeding a few hundred erg/s/cm$^2$.

@@

\begin{figure}
    \script{inhabited_aafo_nuv_thresh.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate is 1.
        %Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
        For all samples, the fraction of inhabited planets drops sharply with increasing threshold \gls{NUV} flux due to the combined effects of never receiving sufficient \gls{NUV} flux for abiogenesis or receiving it before entering the \gls{HZ}.}
        \label{fig:inhabited_aafo_nuv_thresh}
    \end{centering}
\end{figure}
Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for a high abiogenesis rate of $f_\mathrm{life} = 1$.
This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
Another effect responsible for this drop is that some planets receive the required \gls{NUV} flux only before entering the \gls{HZ} -- this is especially likely for M~dwarfs.
For the FGK sample, the fraction of inhabited planets drops at lower threshold fluxes than for the M dwarf sample.

@@

The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
We note that relevant mission concepts, such as the Large Interferometer for Exoplanets~\citep[\life,][]{Quanz2022}, include M-dwarf systems among their primary targets~\citep{Kammerer2018,Carrion-Gonzalez2023}.
Our findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.

@@

Given the challenging nature of detecting and characterizing small (Earth-sized) exoplanets, most exoplanet mission concepts currently under development or considered lack the potential for characterizing large enough samples.
Ground-based 25--40-meter class extremely large telescopes are expected to have the capabilities to detect biosignatures on exoplanets like Proxima Centauri~b~\citep[e.g.,][]{Wang2017,Hawker2019,Zhang2024,Vaughan2024}.
\citet{Hardegree-Ullman2024} used Bioverse to determine potential yields for a 10-year direct imaging and high-resolution spectroscopy survey of O$_2$ on the Giant Magellan Telescope (GMT) and on the Extremely Large Telescope (ELT) and found that between 7 and 19 habitable zone Earth-sized planets could be probed for Earth-like oxygen levels.
Such a sample is too small to test the UV Threshold Hypothesis, but it may be synergistic with other detection methods.

%Meaningful tests of the UV~Threshold Hypothesis may thus require a different survey strategy that increases the sample of characterized planets.
The \gls{HWO} is expected to characterize a sample of $\sim 25$ Earth analogs~\citep{Mamajek2023,Tuchow2024}.
Depending on the technical design, \life\ is expected to target 25–80 \glspl{EEC}~\citep{Kammerer2018,Quanz2022}, which could be just sufficient to constrain the UV~Threshold Hypothesis.
One current exception is the Nautilus Space Observatory concept~\citep{Apai2019,Apai2022}.
Nautilus aims to characterize up to $\sim 1000$ \gls{EEC} via transmission spectroscopy, building on an innovative optical technology.
To guide the definition of future biosignature surveys, it is important to refine predictions on the role of UV radiation in prebiotic chemistry with both theoretical and experimental work.
